Resilience and Home-Space for WSTS \thanks{This work was partly done while the authors were supported by the Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028).}

### 2023-10-20 VMCAI-revised-version/final.tex
corrigé une typi
effectivity

\begin{abstract}
\noindent
Resilience of unperfect systems is a key property for improving safety by insuring that if a system could go into a bad state then it can also leave this bad state and reach a safe state.
We consider six types of resilience (one of them is the home-space property) defined by an upward closed set $\Safe$  ($\Safe=\mathop{\uparrow} \Safe$) or a downward-closed set ($\Safe=\mathop{\downarrow} \Safe$), and by the existence of a bound $k$ on the length of minimal runs starting from $\Bad$ and reaching $\Safe$, where $\Bad$ is generally the complementary of $\Safe$. We study the decidability of each type of resilience for two kinds of models: WSTS and VASS. We first show that most of all resilience problems are undecidable for WSTS with strong compatibility (and with effective $\pred$-basis). Then we prove the decidability of resilience for completion-$\post$-effective $\omega^2$-WSTS with strong compatibility (almost all known WSTS are in this class) and $\Safe = \mathop{\uparrow} \Safe$. Moreover, some resilience properties are decidable for three other classes of WSTS : (1) WSTS with effective 
$\mathop{\uparrow}$ $\post^*$ basis and $\Safe=\mathop{\uparrow} \Safe$ ; (2) ideal-effective WSTS with downward and upward compatibilities ; and (3) ideal-effective downward-compatible WSTS with $\Safe=\mathop{\downarrow} \Safe$. Finally, we study the resilience for VASS with semi-linear subsets $\Bad$ and $\Safe$ and for variations of VASS (lossy counter machines,  $\mathbb{Z}-$VASS and continuous VASS); most of the resilience properties are shown decidable.
\end{abstract}

@@

\mathieu{Donner des exemples dans cette classe.}

>>>>

\begin{abstract}
\noindent
Resilience of unperfect systems is a key property for improving safety by insuring that if a system could go into a bad state in $\Bad$ then it can also leave this bad state and reach a safe state in $\Safe$.
We consider six types of resilience (one of them is the home-space property) defined by an upward-closed set or a downward-closed set $\Safe$, and by the existence of a bound on the length of minimal runs starting from a set $\Bad$ and reaching $\Safe$ ($\Bad$ is generally the complementary of $\Safe$). We study the decidability of each type of resilience for many models: Well Behaved Transition Systems (WBTS), WSTS, VASS and other variations of counter machines. 
%
We first show that all resilience problems are undecidable for effective WSTS with strong compatibility and for upward-closed sets $\Safe$; resilience is also undecidable for reset-VASS (that are WSTS with effective $\pred$-basis) for downward-closed sets $\Safe$.

We then show that resilience is decidable for WBTS with the computability of the finite basis of the downward-closure of the set of (immediate) successors of a downward-closed set ($\post$-basis-effective) and for WSTS with the (well-known) $\pred$-basis-effective hypothesis.
%
%		most of all resilience problems are decidable for WSTS with the good effective hypothesis. R
%
Most of the resilience properties are shown decidable for other classes like downward-WSTS, VASS, lossy counter machines, integer VASS and continuous VASS.

%Then we prove the decidability of resilience for WBTS and WSTS with strong compatibility, adapted effective hypothesis and for upward-closed sets $\Safe$. Moreover, some resilience properties are decidable for three other classes of WSTS : (1) WSTS with effective 
%$\mathop{\uparrow}$ $\post^*$ basis and $\Safe=\mathop{\uparrow} \Safe$ ; (2) ideal-effective WSTS with downward and upward compatibilities ; and (3) ideal-effective downward-compatible WSTS with $\Safe=\mathop{\downarrow} \Safe$. 
%Finally, we study the resilience for VASS with semi-linear subsets $\Bad$ and $\Safe$ and for variations of VASS (lossy counter machines,  $\mathbb{Z}-$VASS and continuous VASS)
\end{abstract}

@@

\mathieu{Donner des exemples dans cette classe.}
\alain{je pense aux LCM avec insertions, elle  devrait avoir toutes les propriétés mais faut vérifier}
